Ultra-compact, widely bandwidth-tunable silicon filter with an unlimited free-spectral range

### 2015-10-21 ReviewersReply.tex

\title{Reply to reviewers}
\author{Jonathan St-Yves}

\begin{document}

\maketitle

\section*{Reviewer 1}
The manuscript entitled "Ultra-compact, widely bandwidth-tunable silicon filter with an unlimited free-spectral range" presents experimental results on an integrated tunable optical filter in the C-band. The manuscript highlights some features of the devices such as the wide tunability (670 GHz), a high-rejection (55dB), the small footprint, and a low group delay variation.
The device is based on a cascaded contra-directional coupler structure, and the authors also show how a four-stage cascaded device could improve some of its characteristics, such as an expanded bandwidth.

The article appears relevant to the scope of Optics Letters and to the research community. The quality of the science and the presentation are very high.
My main concern with this manuscript is with the title, since it suggests "unlimited free-spectral range" where the small print limits that range to the C-band. Since this is misleading, I would rather suggest a more discreet approach.

\emph{The free spectral range is in fact unlimited, at least in the band where silicon and silicon oxide is transparent. The second order reflection theoretically happens at $\lambda_c/2$, or about 770 nm in this case. 
To clarify what we mean by free spectral range: it is the spacing between two successive dropped optical intensity maxima.}

Other than that, here are some questions that, if addressed, might improve the overall quality further:
\begin{itemize}
\item There is some minor inconsistency with the metrics, highlighting a -20dB level requirement (in Figure 3), and then down-grading it to -15dB (in figure 5 - in particular for the +70K curve)

\emph{While -20 dB is the typical standard, -15 dB also has similar practical uses. In cases where side-lobes suppression is important, we expect better performance from the 4 staged design of Figure 7. Changed the text to make it clearer and not possibly imply that -15 is the standard.}

@@

\item Related to the previous point, the shape of the spectral response degrades substantially - beyond acceptable maybe - when the device is pushed to the limit, the performance in this case is arguably consistent with the advertised metrics (top-hat for instance).

\emph{As discussed above, the acceptability of the shape depends of the application. Added the 15 dB figure of merit to the comparison table.}

@@

\item Power dissipation: The device described in the manuscript appears to require a significant amount of power which could cause problems if integrated with electronics as suggested. This also will potentially affect the performance of a device comprising even more stages - increasing the effect of cross-talk. Please comment on that and include an estimation of the performance limits of such enhanced devices to their integration potential.

\emph{Our experiment used a very simplistic heater without thermal isolation or any kind of optimization. Using techniques such as the ones described in the reference of this section, these would not be issues. In principle, we could only dissipate energy through the electrical connections and the optical inputs and outputs, with the rest of the device being suspended in the air and thus thermally isolated. Made some changes to the text to better comment on the issue.}

@@

\emph{Agreed, changed to micrometers squared.}

\item A typo on the 1st paragraph of page 3, "then"; should read "than".

\emph{Thank you for the correction.}

\emph{And lastly, thank you for the comments and the publication recommendation.}

@@

\end{itemize}

\section*{Reviewer 2:}
This manuscript reports on the design and measurement of a tunable filter, composed of a pair of cascaded contra-directional couplers, on a silicon-photonics platform. The overall work, and especially the presentation, of the authors is to be highly commended. Every step of the design and measurement processes is described in succinct but complete detail. The cited references seem to provide adequate sources for further detail. The device is to be applauded for its simplistic design yet impressive filtering capability and tunability. Although the manuscript surpasses the standards of many of its peers, we hope that the authors consider the comments below to further strengthen their submission.

@@

\begin{itemize}
\item A brief discussion on the geometry and tolerance requirements of individual corrugations would be appreciated. If precise geometry is not as important as control of periodicity, for example, then this device could be more properly “CMOS compatible” in that highly precise (and less manufacturing-friendly) electron beam lithography may not be an absolute necessity.

\emph{Large coupling coefficients and precise apodization profiles are easier to achieve with this process, although optical lithography able to resolve the required periodicity and gap between waveguides as seen in Fig.~2~a) could be used. The corrugations do not need to be square shaped, but must have a large enough amplitude.}

\emph{Our experience with optical lithography shows that the design needs to be biased to account for amplitude distortion, but it is in principle possible to use CMOS compatible process, albeit with an additional iteration to obtain the biasing information.}

\item A brief discussion on the coupling coefficients and \textbf{radiation losses} of the contra-directional couplers may be appropriate. We assume that the coefficients play a large role in obtaining the filter shapes present.

\emph{High coupling coefficients are necessary to design short devices with low through port losses and large bandwidth. A comparison with the fabricated device versus a larger coupling is illustrated in Fig. 7.}

@@

\item It is strange that no mention is given of the “through” and “residue” responses (referring to the nomenclature of Fig. 1).

\emph{We choose to focus on the drop port response in this article. INCLUDE THROUGH PORT IN FIG 3?}

@@

\item Following from the foregoing two points, we notice a 2-dB insertion loss in Fig. 3. Is this a result of radiation loss or large amounts of signal exiting from the “through” and “residue” ports?

\emph{Fig. 3 shows an insertion loss of 0.5 dB, both in experiments and in simulation. This loss is  mainly due to silicon absorption and waveguide radiative losses (0.4 dB). The signal exiting from other ports causes about 0.1 dB of losses.}

\item We encourage the authors to drop the term “ultra-compact” from their title. By the authors’ own admission, the device is at least 312 µm long, with at least several microns of width. These dimensions may be compact in terms of conventional optical filters, but in a silicon platform, which tends to promise large-scale integrability of components, these dimensions are far from compact.

\emph{The device ($\approx$ 7000 \textmu m$^2$) is slightly larger than micro-rings ($\approx$ 3500 \textmu m$^2$) and smaller than most MZI designs($\approx$ 600000 \textmu m$^2$) with similar performance. TO DISCUSS WITH WEI }
\end{itemize}

### 2015-10-21 CacadedTunableFilter.tex
Made reviewer changes

\usepackage{ulem}
\usepackage{amsmath,amssymb}
\newcommand{\me}{\mathrm{e}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}

@@

\title{Ultra-compact, widely bandwidth-tunable filter using cascaded silicon photonic contra-directional couplers}

@@

\begin{abstract}
We demonstrate an integrated tunable band-pass filter continuously tuned from 788 GHz down to 117 GHz of bandwidth with up to 3 nm of central wavelength tunability in the infrared C-band. The out-of-band contrast is up to 55 dB, the in-band ripples are under 0.3 dB and the free spectral range is unlimited. This result was achieved using contra-directional couplers and micro-heaters on the silicon-on-insulator platform with a footprint of less than 0.007 mm$^2$. An other iteration pushing the tunability from 1 THz to 100 GHz is discussed. 
\end{abstract}

@@

Next-generation optical transmission systems applying flexible networking and the super-channel technique will require highly dynamic channel allocation to drastically increase the spectral efficiency and transmission capacity \cite{jinno2009spectrum, geisler2011demonstration}.
Tunable optical filters, reconfigurable in both center wavelength and bandwidth with scalability towards terahertz \cite{geisler2011demonstration}, are essential for these applications.
Such large bandwidth tunability is currently only available in bulky bench-top systems using diffractive grating spectrometers or liquid crystals.
Integrated solutions are desired for lower cost and power consumption.
In particular, silicon photonics  based on the sub-micron silicon-on-insulator platform allows for CMOS compatible mass fabrication, enabling low cost, high yield, and high-density chip-scale integration.
Existing solutions for tunable  filters on silicon include devices based on microring resonators \cite{DynamicBW, ong2013ultra} and Mach-Zehnder interferometers (MZIs). These devices have relatively small tunable bandwidth (less than 200 GHz) and small free spectral range (FSR) (typically less than 10 nm), not suitable for high-capacity transmission applications.

@@

The schematic of the proposed device is shown in Fig.~\ref{fig:schematic}. 
It consists of a pair of cascaded contra-DCs, each operating as a drop filter. 
The drop port of the first contra-DC is connected to the input port of the second contra-DC. 
Therefore, the finally dropped signal is determined by the product of the  drop-port transfer functions of the two contra-DCs.

The center wavelength of a contra-DC is determined by the phase-match condition: $\lambda_\text{c} = \Lambda (n_\text{1}+n_\text{2})$, where $\lambda_\text{c}$ is the center wavelength of contra-directional coupling, $\Lambda$ is the grating pitch, and $n_\text{1}$ and $n_\text{2}$ are the effective refractive indices of the first-order and second-order eigenmodes in the coupler. 
At first, the two filters are identical and drop a large well defined window. By changing the temperature of a single filter, the bandwidth of the finally dropped signal can be adjusted by detuning their central wavelengths.

@@

A broadband filter design for a coarse WDM on the 220-nm SOI platform \cite{shi2013siliconCWDM} is adopted for individual contra-DCs.  
In each contra-DC, the widths of the waveguides are of 560 and 440 nm.
This high asymmetry ensures a negligible forward cross-coupling. 
The gratings are formed by sidewall corrugations in strip waveguides with a pitch of 312 nm. 
The spacing between the waveguides varies between 65 and 135 nm to produce a Gaussian coupling profile apodization as shown in Fig.~\ref{fig:SEM} for efficient sidelobe suppression, for which details can be found in \cite{shi2013siliconCWDM}.

The spectral responses of each contra-DC is calculated using coupled mode theory, following the same procedure as in  \cite{shi2013siliconContraDC}, with three additions: apodization, temperature and noise simulation.

While an unapodized (uniform) grating can be simulated with only two transfer matrices , $S_1$ and $S_2$ \cite{shi2013siliconContraDC}, an apodized one needs both of these matrices for each step along the propagation, to match the change of coupling. 
The first change is thus to modify the transfer matrix of the system to the product of all the grating elements along the propagation axis.
Our simulations use 100 segments to approximate the constantly changing coupling.

On top of each contra-DC sits a metal strip acting as a micro-heater for thermal tuning.  
The refractive index of silicon has a temperature dependence of $\frac{\diff{n_\text{Si}}}{\diff{T}}=$ 1.87$\times10^{-4}$K$^{-1}$ at room temperature for wavelengths around 1550 nm~\cite{frey2006temperature}.
Without electrical input, the optical responses of the two filters are identical (neglecting fabrication uniformness) resulting in a sharp transition from the pass-band to the stop-band. 
Applying independent electrical currents on the heaters, we can shift the spectra of the contra-DCs simultaneously or differentially for wavelength or bandwidth tuning.
Ideally, each contra-DC should be uniformly heated so that the apodization profile isn't disturbed as the temperature varies.

In simulation, the temperature dependence is accounted for by increasing the effective index according to 
$\diff{n_\text{eff}}=\frac{\diff{n_\text{Si}}}{\diff{T}}*\frac{\diff{n_\text{eff}}}{\diff{n_\text{Si}}}*\diff{T}$, where $\frac{\diff{n_\text{eff}}}{\diff{n_\text{Si}}}$ is evaluated using an eigenmode solver.

While only coupling apodization has been discussed so far, it is common in Bragg grating profiles to also adjust the local Bragg wavelength. This is done with a phase shift changing the local effective index or period. This design only uses coupling apodization in principle, but phase noise also appears during fabrication due to non-uniformity and sidewall roughness.

There are 2 main types of waveguide noise. High frequency roughness(typically around 10 $\mu$m of spacial period\cite{simard2013characterization}) is caused by sidewall roughness and decreases the sidelobe suppression. Low frequency variation(larger than the device or 300 $\mu$m) is caused by wafer thickness and etching variance and causes a linear chirp that increases the bandwidth.

High frequency phase noise is added to the transfer matrix solution by adding random noise to the $\Delta\hat{\beta}$ propagation constants. The frequency distribution of the noise will thus match the quantization steps of the matrices, but since the response noise is dependent of both the amplitude and frequency of the sidewall roughness \cite{simard2011impact},we can fit the simulation to the experiment by choosing the right noise amplitude.
Similarly, we can also add a variation to $\Delta\hat{\beta}$ which is dependent on the longitudinal position.

@@

The device was fabricated using a CMOS-compatible technology with electron-beam lithography. 
Fiber grating couplers\cite{zhong2014focusingFGC} were used as optical IOs in the measurement. 
Fig.~\ref{fig:passive} shows the measured drop port response of the cascaded contra-DC filters. 
The measurements were normalized using the response of a pair of directly connected fiber grating couplers on the same chip.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Passive62}
\caption{ Spectral response of the cascaded filters without heating: 1-dB BW is 733 GHz, 3-dB BW is 788 GHz, 10-dB BW is 868 GHz, and 20-dB BW is 990 GHz. Inset shows the filter shape up to -20 dB, which is the typical communication requirement.}
\label{fig:passive}
\end{figure}

The device exhibits a high side-lobe suppression ratio (SLSR) of over 40 dB and a high contrast of about 55 dB between the pass-band and the stop-band. 
The insertion loss is very low, less than 0.5 dB (i.e., 0.25 dB per contra-DC), with small ripples of less than 0.3 dB within the 1-dB passband over 5.8 nm (733~GHz). 
The edge roll-off rate is 19 dB/nm on the left side and 24 dB/nm on the right side.
Further work could optimize the grating structure using the inverse layer peeling algorithm\cite{skaar2001synthesis} since the mathematics of the contra-directional coupler are similar to those of Bragg gratings.

By changing the temperature of a single contra-DC, we offset the phase-match condition of a filter, resulting in a smaller band overlap between the two contra-DCs and thus a narrower passband in the drop port, as shown in Fig.~\ref{fig:bandTune}.  
Due to the wavelength detuning,  the stop-band edges are only determined by the single filters. 
As a result, the side-lobes suppression degrades  for small bandwidths but is still better than 15 dB. A continuous tuning of the 3-dB bandwidth from 788 GHz down to 117 GHz (i.e., over 670 GHz or 5.4 nm) is experimentally observed as the on-chip temperature increases by 70 degrees. 
The smallest bandwidth measured in this case was limited by the maximal power delivered before the metal heaters (300-nm-thick, 2-$\mu$m-wide Al strips) were damaged. 
The power efficiency of the bandwidth tuning is about 24 mW/nm, which can significantly improved by optimizing the heater design, e.g., using smaller heater features and thermal isolation \cite{dong2010thermally}.

We observe that the central wavelength at higher temperature is out of the original band, indicating thermal coupling between the two filters. Also, the side-lobes suppression ratio is lower for smaller bandwidth due to the signal only being affected by one filter. To solve both of these problems, next generation devices should use a isolation or larger spacing between filter elements (we used 12 \text{$\mu$}m center to center).

\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Band6}
\caption{Spectral response of the device for different temperatures applied to only one contra-DC: 1-dB BW tuned down to 65 GHz, 3-dB BW to 117 GHz. The temperatures are obtained from simulation fit. Inset shows the temperature dependence.}
\label{fig:bandTune}
\end{figure} 
An improvement upon this design is simulated in Fig.~\ref{fig:bandTuneSimu}. By using 4 filtering elements instead of 2, and improving the thermal performances to achieve a stable high temperature, it would be possible to tune the bandwidth from over 1 THz down to under 100 GHz while still maintaining a very high out of band contrast of over 20 dB. This simulation uses the noise parameters observed in the 2 stage experiment.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/4StageDesignCompiled}
\caption{Simulation of the spectral response with the heat applied to two out of four cascaded contra-DCs: the bandwidth can be continuously tuned from 1020 GHz to  51GHz.}
\label{fig:bandTuneSimu}
\end{figure}

By applying the same temperature variation on both contra-DCs, the center wavelength can be tuned without affecting the filter shape.
As shown in Fig.~\ref{fig:wavTune}, when the center wavelength is continually changed over 4 nm by varying the on-chip temperature, the filter shape is maintained with sharp edges. Actually, slight detuning between the cascaded contra-DCs may be used to compensate for band-edge distortions due to fabrication errors for a more symmetric filter shape. 
The power efficiency of the wavelength tuning is about 44 mW/nm, which is about twice the power consumption of the bandwidth tuning since two contra-DCs are heated simultaneously.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Central2}
\caption{Spectral response with the heat applied to both contra-DCs: the central wavelength is tuned from 1535 nm to 1539 nm continuously; the temperatures are calculated by comparison to simulation. Inset shows the temperature dependence.}
\label{fig:wavTune}
\end{figure}

%\subsection{Phase/dispertion}
An other important parameter to monitor is the group delay of the device. An uneven delay can cause distortion during detection of the signal, just like dispersion. The simulation and measurements seen in Fig.~\ref{fig:phase} both show a group delay difference of less than 10 ps.
\begin{figure}[tbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Phase2}
\caption{ Group delay response of the tunable filter at the drop port. The delay stays within 12 ps between any two points in the band. The out of band results are noisy due to the weak signal.}
\label{fig:phase}
\end{figure}

Table \ref{table:comparison} shows a comparison of this filter with other recent publications of tunable filters. No other tunable filter has such a large available bandwidth and FSR, making contra-DC unique in their application potential.

%\section{Conclusion}
In summary, we have demonstrated a bandwidth tunable filter with a low insertion loss of less than 0.5 dB, low ripples of less than 0.3 dB, a large maximal bandwidth of greater than 750 GHz, and high contrast of 55 dB. 
A large bandwidth tuning range over 670 GHz has been achieved, which, to the best of our knowledge, is the widest ever demonstrated on a silicon chip. 
This ultra-wide bandwidth tunability makes the device very attractive for next-generation ultrahigh baud-rate applications (e.g., high-capacity super-channel transmissions) and flexible optical networking.

\begin{table*}[thb]
\caption{Recent results with multi-element on-chip silicon filters}
\begin{tabular}{cccccc}
    \hline
	Publication & Filter type & BWmax / FSR & Tunable BW & On-chip loss & Contrast \\
    \hline
    %Ibrahim (2011) &	Cascaded MRs &	0.4 GHz / 10 GHz &	0.6-2 GHz &	0.6 dB &	30 dB
    % \\
    %Alipour (2011) &	Cascaded MRs &	5 GHz / 650 Ghz &	0.8-5 GHz &	1.25-3.75 dB &	38 dB
    % \\
    Ding (2011)\cite{ding2011bandwidth} &	MRs + MZI &	55 GHz / 1 THz &	28-55 GHz &	3.6 dB &	30 dB
     \\
 	Orlandi (2012)\cite{orlandi2012reconfigurable} &	MRs + MZI &	173 GHz / 200 GHz &	23-173 GHz &	0.46-1.06 dB &	15-34 dB
      \\
    %Luo (2012) &	Cascaded MRs &	100 GHz / 750 Ghz &	No &	0.3-0.6 dB &	50 dB  \\
    Ong (2013)\cite{ong2013ultra} &	Cascaded MRs &	125 GHz/ 0.9 THz &	11.6-125 GHz &	0.25 dB &	50-100 dB
      \\
    \textbf{This experiment }& Cascaded contra-DC &	778 GHz / Unlimited  &	117-778 GHz &	0.5 dB &	55 dB 
    \\
    \textbf{4 Stages }& Cascaded contra-DC &	1020 GHz / Unlimited  &	51-1020 GHz &	<1 dB &	>55 dB 
    \\

@@

\section*{Acknowledgments}
We acknowledge CMC Microsystems for the  software and the fabrication subsidy. We thank Yun Wang and Lukas Chrostowski with the Univeristy of British Columbia for the fiber grating coupler design. The authors acknowledge the Natural Sciences and Engineering Research Council of Canada, TeraXion and Prompt for funding this research. The silicon chip was fabricated at the University of Washington Microfabrication/Nanotechnology User Facility, a member of the NSF National Nanotechnology Infrastructure Network.

>>>>

\usepackage{ulem}
\usepackage{amsmath,amssymb}

@@

\title{Ultra-compact, widely bandwidth-tunable silicon filter with an unlimited free-spectral range}

@@

\begin{abstract}
Next-generation high-capacity optical networks require flexible allocation of spectrum resources, for which low-cost optical filters with an ultrawide bandwidth tunability beyond hundred GHz are desired.
We demonstrate an integrated band-pass filter with the bandwidth continuously tuned across 670 GHz (117 GHz to 788 GHz), which, to the best of our knowledge, is the widest tuning span ever demonstrated on a silicon chip.
The filter also features simultaneous wavelength tuning and an unlimited free spectral range. 
We measured an out-of-band contrast of up to 55 dB, low in-band ripples of less than 0.3 dB, and in-band group delay variation of less than 8 ps.
This result was achieved using cascaded Bragg-grating-assisted contra-directional couplers and micro-heaters on the 220-nm silicon-on-insulator platform with a footprint of less than 7000~\textmu m$^2$. Another design with the bandwidth continuously tunable from 50 GHz to 1 THz is also presented. 
\end{abstract}

@@

Next-generation optical transmission systems applying flexible networking and the super-channel technique will require highly dynamic channel allocation to drastically increase the spectral efficiency and transmission capacity \cite{jinno2009spectrum, geisler2011demonstration}.
Tunable optical filters, reconfigurable in both center wavelength and bandwidth with scalability towards terahertz \cite{geisler2011demonstration}, are essential for these applications.
Such large bandwidth tunability is currently only available in bulky bench-top systems using diffractive grating spectrometers or liquid crystals.
Integrated solutions are desired for lower cost and power consumption.
In particular, silicon photonics  based on the sub-micron silicon-on-insulator platform allows for CMOS compatible mass fabrication, enabling low cost, high yield, and high-density chip-scale integration.
Existing solutions for tunable  filters on silicon include devices based on microring resonators \cite{DynamicBW, ong2013ultra} and Mach-Zehnder interferometers (MZIs). These devices have relatively small tunable bandwidth (less than 200 GHz) and small free spectral range (FSR) -- typically less than 10 nm, unable to cover the entire C-band, which are not suitable for high-capacity transmission applications.

@@

The schematic of the proposed device is shown in Fig.~\ref{fig:schematic}. 
It consists of a pair of cascaded contra-DCs, each operating as a drop filter. 
The drop port of the first contra-DC is connected to the input port of the second contra-DC. 
Therefore, the finally dropped signal("Drop port of the system"), is determined by the product of the  drop-port transfer functions of the two contra-DCs.

%The center wavelength of the contra-DC is determined by the phase-match condition: $\lambda_\text{c} = \Lambda (n_\text{1}+n_\text{2})$, where $\Lambda$ is the grating pitch, and $n_\text{1}$ and $n_\text{2}$ are the effective indices of the first-order and second-order eigenmodes in the coupler. 
At first, the two filters are identical and drop a large, well-defined window. By changing the temperature of a single filter, the bandwidth of the finally dropped signal can be adjusted by detuning the center wavelengths of the two contra-DCs.

@@

A broadband filter design for a coarse WDM on the 220-nm SOI platform \cite{shi2013siliconCWDM} is adopted for individual contra-DCs.  
In each contra-DC, the widths of the waveguides are 560 nm and 440 nm.
This high asymmetry ensures a negligible forward cross-coupling. 
The gratings are formed by sidewall corrugations in strip waveguides with a pitch of 312 nm.
Each contra-DC has 1,000 grating elements for a length of 312 \textmu m.
The spacing between the waveguides varies between 65 and 135 nm to produce a Gaussian-profiled coupling apodization as shown in Fig.~\ref{fig:SEM} for efficient sidelobe suppression, for which details can be found in \cite{shi2013siliconCWDM}.

The spectral responses of each contra-DC is calculated using coupled mode theory and the transfer matrix method, following the same procedure as in  \cite{shi2013siliconContraDC}, with three additions: apodization, temperature dependence and noise simulation.
While an unapodized (uniform) grating can be simulated with only two transfer matrices,
an apodized design needs an independent matrix for each grating element along the propagation, to match the change of coupling, and the transfer matrix of the entire contra-DC is given by the product of all the grating elements along the propagation axis.
Our simulations use 100 segments to approximate the continuously varied coupling.

On top of each contra-DC sits a metal strip acting as a micro-heater for thermal tuning.  
The refractive index of silicon has a temperature dependence of 
${\mathrm{d}{n_\text{Si}}}/{\mathrm{d}T}=$ 1.87$\times10^{-4}$K$^{-1}$
at room temperature for wavelengths around 1550 nm~\cite{frey2006temperature}.
Without electrical input, the optical responses of the two filters are in principle identical (except for fabrication errors), resulting in a sharp transition from the pass-band to the stop-band. 
Applying independent electrical currents on the heaters, we can shift the spectra of the contra-DCs simultaneously or differentially for wavelength or bandwidth tuning.
Ideally, each contra-DC should be uniformly heated so that the apodization profile would not be disturbed as the temperature varies.
In simulation, the temperature dependence is accounted for by 
$\mathrm{d}n_\text{eff}=\mathrm{d}n_\text{Si}/{\mathrm{d}T}\times\mathrm{d}n_\text{eff}/{\mathrm{d}n_\text{Si}}\times\mathrm{d}T$, where $\mathrm{d}n_\text{eff}/{\mathrm{d}n_\text{Si}}$ is evaluated using an eigenmode solver.

While only coupling apodization has been discussed so far, it is common in Bragg grating profiles to also adjust the local Bragg wavelength. This can be done with a phase shift changing the local effective index or period. This design only uses coupling apodization in principle, but phase noise also appears during fabrication due to non-uniformity and sidewall roughness.

There are two main types of waveguide noise considered in our simulation. High-frequency phase noise (typically around 10 \textmu m of spacial period \cite{simard2013characterization}) is caused by sidewall roughness and decreases the sidelobe suppression. Low-frequency phase noise, with a correlation length larger than the device (300 \textmu m), is caused by the wafer thickness non-uniformity and etching variance, resulting in a linear chirp that increases the bandwidth.
High frequency phase noise is considered in our simulation by adding random noise $\Delta\hat{\beta}$ to the propagation constants. The frequency distribution of the noise thus matches the quantization steps of the matrices, but since the response noise is dependent on both the amplitude and frequency of the sidewall roughness \cite{simard2011impact}, we can fit the simulation to the experiment by choosing the right noise amplitude.
Similarly, we can also add a variation to $\Delta\hat{\beta}$ which is dependent on the longitudinal position.

@@

The device was fabricated using a CMOS-compatible technology with electron-beam lithography. 
Large coupling coefficients and precise apodization profiles are easier to achieve with this process, although optical lithography able to resolve the required periodicity and gap between waveguides as seen in \ref{fig:SEM} a) could be used. The corrugations do not need to be square shaped, but must have a large enough amplitude
.
Fiber grating couplers \cite{zhong2014focusingFGC} are used as optical inputs and outputs in the measurement. 
Fig.~\ref{fig:passive} shows the measured drop port response of the cascaded contra-DC filters. 
The measurements were normalized using the response of a pair of directly connected fiber grating couplers on the same chip.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Passive62}
\caption{ Spectral response of the cascaded filters without heating: 1-dB BW is 733 GHz, 3-dB BW is 788 GHz, 10-dB BW is 868 GHz, and 20-dB BW is 990 GHz. Inset shows the filter shape up to -20 dB, which is the typical communication requirement.}
\label{fig:passive}
\end{figure}

The device exhibits a high side-lobe suppression ratio (SLSR) of over 40 dB and a high contrast of about 55 dB between the pass-band and the stop-band. 
The insertion loss is very low, less than 0.5 dB (i.e., 0.25 dB per contra-DC), with small ripples of less than 0.3 dB within the 1-dB passband over 5.8 nm (733~GHz). 
The edge roll-off rate is 19 dB/nm on the left side and 24 dB/nm on the right side.
The simulation agrees well with experiment.

%\subsection{Phase}
\begin{figure}[tbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Phase3}
\caption{ Group delay response of the tunable filter at the drop port. The delay stays within 8 ps between any two points in the band. The out of band results are noisy due to the weak signal.}
\label{fig:phase}
\end{figure}
Another important parameter to monitor is the group delay, as an uneven delay within the pass band may cause signal distortion. The simulation and measurements seen in Fig.~\ref{fig:phase} both show a group delay difference of less than 8~ps within the 1-dB bandwidth. 
Compared to the original design without considering the phase noise due to fabrication errors, the measured group delay shows a slight distortion. Good agreement is obtained by taking into consideration the linear chirp caused by the silicon wafer thickness variation.

Further work could optimize the grating structure further using the inverse layer peeling algorithm\cite{skaar2001synthesis} since the mathematics of the contra-directional coupler are similar to those of Bragg gratings.

By changing the temperature of a single contra-DC, we shift its phase-match condition, resulting in a smaller band overlap between the two contra-DCs and thus a narrower passband in the drop port, as shown in Fig.~\ref{fig:bandTune}.  
Due to the wavelength detuning,  the stop-band edges are now determined by the single filters. 
As a result, the side-lobes suppression for small bandwidths degrades to -15 dB.
A continuous tuning of the 3-dB bandwidth from 788 GHz down to 117 GHz (i.e., over 670 GHz or 5.4 nm) is experimentally observed as the on-chip temperature increases by 70 degrees.

The microheaters are made of 300-nm-thick, 2-\textmu m-wide Al strips.
The power efficiency of the bandwidth tuning is about 24~mW/nm, which dissipates a lot of heat and would be impractical for integration with electronics.
We also observe a slight red-shift of the center wavelength at higher temperature, indicating thermal coupling between the two contra-DCs that are about 10 \textmu m away from each other.
These issues can be resolved by optimizing the heater design, e.g., using smaller heater features and thermal isolation \cite{dong2010thermally}.

\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Band6}
\caption{Spectral response of the device for different temperatures applied to only one contra-DC: 1-dB BW tuned down to 65 GHz, 3-dB BW to 117 GHz. The temperatures are obtained from simulation fit. Inset shows the temperature dependence.}
\label{fig:bandTune}
\end{figure}

By applying the same temperature variation on both contra-DCs, the center wavelength can be tuned without affecting the filter shape.
As shown in Fig.~\ref{fig:wavTune}, when the center wavelength is continually changed over 4 nm by varying the on-chip temperature, the filter shape is maintained with sharp edges. 
%Actually, slight detuning between the cascaded contra-DCs may be used to compensate for band-edge distortions due to fabrication errors for a more symmetric filter shape. 
The power efficiency of the wavelength tuning is about 44~mW/nm, which is about twice the power consumption of the bandwidth tuning since two contra-DCs are heated simultaneously.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Central2}
\caption{Spectral response with the heat applied to both contra-DCs: the central wavelength is tuned from 1535 nm to 1539 nm continuously; the temperatures are calculated by comparison to simulation. Inset shows the temperature dependence.}
\label{fig:wavTune}
\end{figure}

\begin{table*}[thb]
\caption{Recent results with multi-element on-chip silicon filters}
\begin{tabular}{cccccc}
\hline
Publication & Filter type & BWmax / FSR & Tunable BW & On-chip loss / Ripples & Contrast \\
\hline
Ding (2011)\cite{ding2011bandwidth} &	MRs + MZI &	55 GHz / 1 THz &	28-55 GHz &	3.6 dB / 1 dB &	30 dB
\\
Orlandi (2012)\cite{orlandi2012reconfigurable} &	MRs + MZI &	173 GHz / 200 GHz &	23-173 GHz &	0.46-1.06 dB / 0.5 dB &	15-34 dB
\\
Ong (2013)\cite{ong2013ultra} &	Cascaded MRs &	125 GHz/ 0.9 THz &	11.6-125 GHz &	0.25 dB / 3 dB &	50-100 dB
\\
\textbf{This experiment }& Cascaded contra-DC &	778 GHz / Unlimited  &	67-778 GHz &	0.5 dB / 0.3 dB &	15-55 dB 
\\
\textbf{4-stage design  }& Cascaded contra-DC &	1 THz / Unlimited  &	49-1007 GHz &	<1 dB / 0.3 dB & >49 dB 
\\

@@

Table \ref{table:comparison} shows a comparison of recent publications of tunable optical filters on the SOI platform. 
Our device shows the highest tunable bandwidth and is the only one that allows for a bandwidth beyond 200~GHz. 
It is noteworthy that the other devices all have a small FSR less than the span of the C-band (35~nm), while our device does not suffer from a limited FSR.
This indicates a unique capacity for dynamic allocation of spectrum resources.
In addition, our device is among the best in terms of loss ripples and contrast.

\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/4StageDesignCompiled2}
\caption{Simulation of the spectral response with the heat applied to two out of four cascaded contra-DCs: the 1 dB bandwidth can be continuously tuned from 1007 GHz to  49GHz. The sidelobes stay under -49 dB even for small bandwidths due to the double filtering. This simulation doesn't consider fabrication phase noise.}
\label{fig:bandTuneSimu}
\end{figure}

For improved performance and flexibility, more stages can be cascaded.
Fig.~\ref{fig:bandTuneSimu} shows our simulation results of a 4-stage cascade contra-DC filter, consisting of two pairs of identical contra-DCs with a stronger coupling than the one experimentally demonstrated (53500~m$^{-1}$ vs 38000~m$^{-1}$). Each pair is controlled by the same temperature, performing as a single tuning element.

%\section{Conclusion}
In summary, we have demonstrated a bandwidth tunable filter with a low insertion loss of less than~0.5 dB, low ripples of less than~0.3 dB, a large maximal bandwidth of greater than 750~GHz, and high contrast of 55~dB. 
A large bandwidth tuning range over 670~GHz has been achieved, which, to the best of our knowledge, is the widest ever demonstrated on a silicon chip. 
This ultra-wide bandwidth tunability makes the device very attractive for next-generation ultrahigh baud-rate applications (e.g., high-capacity super-channel transmissions) and flexible optical networking.

\section*{Acknowledgments}
We acknowledge CMC Microsystems for the  software and the fabrication subsidy. We thank Yun Wang and Lukas Chrostowski with the University of British Columbia for the fiber grating coupler design. The authors acknowledge the Natural Sciences and Engineering Research Council of Canada for funding this research. This work is part of the SPEED research project (Silicon Photonic Electrically Engineered Devices) funded by NSERC (RDCPJ438811-12), PROMPT (PJT-2011-17), and TeraXion. The silicon chip was fabricated at the University of Washington Microfabrication/Nanotechnology User Facility, a member of the NSF National Nanotechnology Infrastructure Network.
